Bayesian Inference Using Sequential Monte-Carlo Algorithm for Dynamic System Models

### 2020-08-09 MSc-dissertation.tex
Bib restructure

\begin{thebibliography}{100}

    \bibitem{axonal} Becker, T., Wullimann, M.F., Becker, C.G., Bernhardt, R.R. and Schachner, M., 1997. Axonal regrowth after spinal cord transection in adult zebrafish. Journal of Comparative Neurology, 377(4), pp.577-595.

    \bibitem{ref:Tsarouchas} Tsarouchas, T.M., Wehner, D., Cavone, L., Munir, T., Keatinge, M., Lambertus, M., Underhill, A., Barrett, T., Kassapis, E., Ogryzko, N. and Feng, Y., 2018. Dynamic control of proinflammatory cytokines Il-1$\beta$ and Tnf-$\alpha$ by macrophages in zebrafish spinal cord regeneration. Nature communications, 9(1), pp.1-17.

    \bibitem{Anderson} Anderson, W.D., Makadia, H.K., Greenhalgh, A.D., Schwaber, J.S., David, S. and Vadigepalli, R., 2015. Computational modeling of cytokine signaling in microglia. Molecular BioSystems, 11(12), pp.3332-3346.

    \bibitem{ref:pyabc} Klinger, E., Rickert, D. and Hasenauer, J., 2018. pyABC: distributed, likelihood-free inference. Bioinformatics, 34(20), pp.3591-3593.

    \bibitem{ref:kernel} Filippi, S., Barnes, C. P., Cornebise, J., and Stumpf, M..H. (2013). On optimality of kernels for approximate Bayesian computation using sequential Monte Carlo. Statistical Applications in Genetics and Molecular Biology 12, 1, 87-107.

    \bibitem{ref:abcsysbio} Liepe, J., Kirk, P., Filippi, S., Toni, T., Barnes, C.P., and Stumpf, M.P.H., 2014. A framework for parameter estimation and model selection from experimental data in systems biology using approximate Bayesian computation. Nature Protocols. 9 (2) pp. 439–456.

    \bibitem{ref:MCMC} Gilks, W.R., Richardson, S. and Spiegelhalter, D.J., 1996. Markov Chain Monte Carlo in Practice Chapman \& Hall: London. Markov Chain Monte Carlo in practice. Chapman and Hall, London.

    \bibitem{ABC_rejection} Tavaré, S., Balding, D.J., Griffiths, R.C. and Donnelly, P., 1997. Inferring coalescence times from DNA sequence data. Genetics, 145(2), pp.505-518.

    \bibitem{summaryD} Joyce, P. and Marjoram, P., 2008. Approximately sufficient statistics and Bayesian computation. Statistical applications in genetics and molecular biology, 7(1).

    \bibitem{summaryD2} Nunes, M.A. and Balding, D.J., 2010. On optimal selection of summary statistics for approximate Bayesian computation. Statistical applications in genetics and molecular biology, 9(1).

    \bibitem{threshold} Silk, D., Filippi, S. and Stumpf, M.P., 2012. Optimizing threshold-schedules for approximate Bayesian computation sequential Monte Carlo samplers: applications to molecular systems. arXiv preprint arXiv:1210.3296.

    \bibitem{population} Klinger, E. and Hasenauer, J., 2017, September. A scheme for adaptive selection of population sizes in approximate Bayesian computation-sequential Monte Carlo. In International Conference on Computational Methods in Systems Biology (pp. 128-144). Springer, Cham.

    \bibitem{model_compare} Toni, T. and Stumpf, M.P., 2010. Simulation-based model selection for dynamical systems in systems and population biology. Bioinformatics, 26(1), pp.104-110.

    \bibitem{cosmology} Weyant, A., Schafer, C. and Wood-Vasey, W.M., 2013. Likelihood-free cosmological inference with type Ia supernovae: approximate Bayesian computation for a complete treatment of uncertainty. The Astrophysical Journal, 764(2), p.116.

    \bibitem{ref:compare} Daly AC, Cooper J, GavaghanDJ, Holmes C. 2017. Comparing two sequentialMonte Carlo samplers for exact andapproximate Bayesian inference on biologicalmodels.J. R. Soc. Interface14: 20170340.

    \bibitem{ref:disease} Minter, A. and Retkute, R., 2019. Approximate Bayesian Computation for infectious disease modelling. Epidemics, 29, p.100368.

    \bibitem{ref:adpt_dis} Prangle, D., 2017. Adapting the ABC distance function. Bayesian Analysis, 12(1), pp.289-309.

    \bibitem{ref:adpt_pop} Klinger, E. and Hasenauer, J., 2017, September. A scheme for adaptive selection of population sizes in approximate Bayesian computation-sequential Monte Carlo. In International Conference on Computational Methods in Systems Biology (pp. 128-144). Springer, Cham.

    \bibitem{Toni} Toni, T., Welch, D., Strelkowa, N., Ipsen, A. and Stumpf, M.P., 2009. Approximate Bayesian computation scheme for parameter inference and model selection in dynamical systems. Journal of the Royal Society Interface, 6(31), pp.187-202.

    \bibitem{sensitivity} Gutenkunst, R.N., Waterfall, J.J., Casey, F.P., Brown, K.S., Myers, C.R. and Sethna, J.P., 2007. Universally sloppy parameter sensitivities in systems biology models. PLoS Comput Biol, 3(10), p.e189.

\end{thebibliography}

>>>>

\bibliographystyle{ieeetr}

\bibliography{bibtex/ref}

### 2020-08-09 background/background.tex
Bib restructure

where $l$ represents the likelihood function. In words, Eqn \ref{eq:bayes2} can be described as `posterior is proportional to the product of likelihood and prior', and here posterior of parameters is our target. Prior $p(\theta)$ should express our initial believes or estimates on the parameter valeus before observing the data and conducting experiments, and likelihood function $l(D|\theta)$ is used to describe how well the parameter set $\theta$ explains the data $D$, in probability numbers. Once the data is given, $\theta$ is the only parameter in the likelihood function. Some optimisation-based methods try to find an estimation of parameter set $\theta$ s.t. maximise the likelihood, such methods will struggle when facing a problem where it is hard to specify an analytical expression of likelihood function, let alone the derivative.

>>>>

where $l$ represents the likelihood function. In words, Eqn \ref{eq:bayes2} can be described as `posterior is proportional to the product of likelihood and prior', and here posterior of parameters is our target. Prior $p(\theta)$ should express our initial believes or estimates on the parameter values before observing the data and conducting experiments, and likelihood function $l(D|\theta)$ is used to describe how well the parameter set $\theta$ explains the data $D$, in probability numbers. Once the data is given, $\theta$ is the only parameter in the likelihood function. Some optimisation-based methods try to find an estimation of parameter set $\theta$ s.t. maximise the likelihood, such methods will struggle when facing a problem where it is hard to specify an analytical expression of likelihood function, let alone the derivative.

### 2020-08-09 implementation/implementation.tex
Bib restructure

According to our preliminary studies, ABC SMC framework implemented using \verb|pyABC| in \verb|Python| was adopted in this project. \verb|pyABC| is a popular SMC software packages \cite{ref:pyabc} that has been used form many SMC studies and applications \cite{ref:adpt_pop}. \verb|pyABC| provides an open-source framework for likelihood-free inference, which is an implementation of ABC SMC algorithm and a tool-box for our own inference and analysis tasks. Besides, it supports multi-core execution implemented using \verb|multiprocessing| package for scaling-up, which is of our interests and suitable for the followed performance experiments.

@@

Adaptive population \cite{ref:adpt_pop} can adapt the population size of each generation according to the mean correlation of variation error of the previous population. In our early test the maximal population size that is allowed is set to 5,000 and the adaptive strategy always select the upper bound, as a results of wide parameter space or the wide posterior approximation shape.

>>>>

According to our preliminary studies, ABC SMC framework implemented using \verb|pyABC| in \verb|Python| was adopted in this project. \verb|pyABC| is a popular SMC software packages \cite{ref:pyabc} that has been used form many SMC studies and applications \cite{population}. \verb|pyABC| provides an open-source framework for likelihood-free inference, which is an implementation of ABC SMC algorithm and a tool-box for our own inference and analysis tasks. Besides, it supports multi-core execution implemented using \verb|multiprocessing| package for scaling-up, which is of our interests and suitable for the followed performance experiments.

@@

Adaptive population \cite{population} can adapt the population size of each generation according to the mean correlation of variation error of the previous population. In our early test the maximal population size that is allowed is set to 5,000 and the adaptive strategy always select the upper bound, as a results of wide parameter space or the wide posterior approximation shape.
